---
title: 'Stat. Inf. II: Assignment 1'
author: 'Naimul Chowdhury'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading:
  a. Read the remaining Sections in Chapter 9: 9.3, 9.5 and 9.6. And, read the final section about errors and power on the R handout for inference about a proportion.
  

## Problem 2:

 1. Write a _prop.sample.size()_ function that will output the sample size needed for a one-sample proportion test to achieve

- a desired margin of error (argument #1)

- for a given confidence level (argument #2)

in the "worst-case scenario" (as was explained in class). What was meant by the "worst-case scenario"?

### Solution 1.

We recall that 

$$m = z_{1 - \alpha/2} \cdot \sqrt{\frac{\hat{p}(1-\hat{p})}{n}}.$$


We wish to compute the desired sample size given margin of error $m$ and confidence level $1- \frac{\alpha}{2}$.

In particular, we wish to account for the "worst-case scenario", where the standard error is maximized by our choice of $\hat{p}$.


```{r}
prop.sample.size <- function(m, conf) {
  n <- (qnorm(conf)^2 * 0.5^2)/(m^2)
  response <- c("The sample size for a one-sample proportion test with a desired margin of error", m, " and given confidence level", conf, " is ", n)
  return(response)
}
```


For example, suppose $m = 0.4$, and the given confidence level is 0.995. Then, 
```{r}
prop.sample.size(0.4, .995)
```


2. use your _prop.sample.size()_ from part 1 to do exercise 8.50 from the Agresti book.

How many businesses fail? A study is planned to estimate the proportion of businesses started in the year 2006 that failed within five years of their start-up. How large a sample size is needed to guarantee estimating this proportion correct to within

a. 0.10 with probability 0.95?

```{r}
prop.sample.size(0.10, 0.95)
```

b. 0.05 with probability 0.95?

```{r}
prop.sample.size(0.05, 0.95)
```

c. 0.05 with probability 0.99?

```{r}
prop.sample.size(0.05, 0.99)
```

d. Compare sample sizes for parts a and b, and b and c, and summarize the effects of decreasing the margin of error and increasing the confidence level.

3. Write a _mean.sample.size()_ function that will output the sample size needed for a one-sample mean test to achieve

- a desired margin of error (argument #1)

- for a given confidence level (argument #2)

- for a given standard deviation (argument #3)

Proceed to use that function in order to do exercise 8.53 from the Agresti book.

### Solution 3

Recall that

$$t_{(n-1), \ 1-\alpha/2} \cdot \frac{s}{\sqrt{n}} = m,$$

implying,

$$n = \frac{ \sigma^2 z^2 } {m^2}.$$

This is working under the assumption that we have a population size of at least $n = 30$, since this would make our distrubution approach the normal distribution.


Thus we can construct our function.

```{r}
mean.sample.size <- function(m, conf, s) {
  n <- ((s * qnorm(conf))/m)^2
  response <- c("The sample size for a one-sample mean test with a desired margin of error,", m, " given confidence level", conf, "and standard deviation", s,  " is ", n)
  return(response)
}
```









