---
title: 'Statistics Assignment #7'
author: "Naeem Chowdhury"
date: "6/16/2020"
output: pdf_document
---


##1. Setup
### options
Set up global options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70))
knitr::opts_chunk$set(fig.height=4, fig.width=6)
```

### libraries
Load in needed libraries 
```{r, include=FALSE}
library(tidyverse)
library(RColorBrewer)
library(ggplot2)
#library(haven)
#library(plotly)
#install.packages('rgl')
#library(rgl)
#library(ISLR)
```

## 2. File management
### Create variables for directories
```{r file_management, include=FALSE}
project.dir <- getwd() #naeem
output.dir <- "/Output"
data.dir <- "C:/Users/Naeem Cho/Desktop/School Work/Statistical Inference/Datasets"
setwd(project.dir)
getwd()
```

## 3. Importing Data
```{r, echo = FALSE, include=FALSE}

```

# Problem #1

  1. See slides #54-55. __Hint:__ At one point in your derivation, make sure to utilize the fact that $D_{fem} = 1$ for females, 0 for males. It will allow you to "break the sum": $\sum_{i = 1}^n = \sum_{i \in Females} + \sum_{i \in Males}$.
  
  2. On slide #60, attempt to interpret the effect of what happens "per 1-unit increase in $D_{Asia}$, while holding other variables constant". Why is it not possible ot evaluate that effec there? Hence, to avoid that, how many dummy variables should we use in order to model a categorical predictor in this case? Show that now, having dropped $D_{Afr-Am}$, we can actually evaluate the aforementioned effect of $D_{Asia}$.
  
  3. When dealing with a categorical predictor $X$ containing > 2 categories (say, 3), why do we have to use dummy variables approach? Why not just model it as
  
  $$Piecewise function here$$.

# Problem #2

For the _Wage_ data set from _ISLR_ package, we will use _age_ and _race_ to predict person's _wage_ (in $1000, I believe).

  1.
    a. Proceed to write down the full modeling equation for $wage \sim age + race$ regression.
    b.Fit the model from 1(a), and write down the fitted equation.
    c. Interpret the most statistically significant __dummy variable__ (NOT the "per 1-unit" version though, the group comparison version).
    d. Conduct the test for significance of _race_ when predicting person's wage. In particular, make sure to 1) formulate the $H_0$, $H_a$ hypotheses; 2) Write down the modeling equation of the "null" model; 3) write the formula for test statistic; 4) use _R_'s _anova()_ to carry out the test, and match the output to the terms in the test statistic formula; 5) Provide the conclusion of the test.

  2. 
    a. Provide data visualization (use your fancy _ggplot2_) that will help determining if there's a strong interaction between _race_ and _age_ when predicting _wage_. Comment on whether you observe an interaction. Explain. __NOTE(!)__: You shouldn't start fitting any models with interactions yet.
    b. Proceed to confirm your hunch from part 2(a) by actually conducting an appropriate statistical test for significance of interaction. In particular, make sure to 1) write the full modeling equation __with interaction__; 2) forumlate the $H_0$, $H_a$ hypotheses; 3) write down the modeling equation of the "null" model; 4) use _R_'s _anova()_ to carry out the test (no need to explain details here); 5) Provide the conclusion of the test.
    
    
# Problem #3
