---
title: 'Statistics Assignment #7'
author: "Naeem Chowdhury"
date: "6/16/2020"
output: pdf_document
---


##1. Setup
### options
Set up global options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70))
knitr::opts_chunk$set(fig.height=4, fig.width=6)
```

### libraries
Load in needed libraries 
```{r, include=FALSE}
library(tidyverse)
library(RColorBrewer)
library(ggplot2)
#library(haven)
#library(plotly)
#install.packages('rgl')
#library(rgl)
#library(ISLR)
```

## 2. File management
### Create variables for directories
```{r file_management, include=FALSE}
project.dir <- getwd() #naeem
output.dir <- "/Output"
data.dir <- "C:/Users/Naeem Cho/Desktop/School Work/Statistical Inference/Datasets"
setwd(project.dir)
getwd()
```

## 3. Importing Data
```{r, echo = FALSE, include=FALSE}

```

# Problem #1

  1. See slides #54-55. __Hint:__ At one point in your derivation, make sure to utilize the fact that $D_{fem} = 1$ for females, 0 for males. It will allow you to "break the sum": $\sum_{i = 1}^n = \sum_{i \in Females} + \sum_{i \in Males}$.
  
  2. On slide #60, attempt to interpret the effect of what happens "per 1-unit increase in $D_{Asia}$, while holding other variables constant". Why is it not possible ot evaluate that effec there? Hence, to avoid that, how many dummy variables should we use in order to model a categorical predictor in this case? Show that now, having dropped $D_{Afr-Am}$, we can actually evaluate the aforementioned effect of $D_{Asia}$.
  
  3. When dealing with a categorical predictor $X$ containing > 2 categories (say, 3), why do we have to use dummy variables approach? Why not just model it as
  
  $$Piecewise function here$$.

# Problem #2

# Problem #3
