---
title: 'Statistics Assignment #10'
author: "Naeem Chowdhury"
date: "6/26/2020"
output: pdf_document
---


##1. Setup
### options
Set up global options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70))
knitr::opts_chunk$set(fig.height=4, fig.width=6)
```

### libraries
Load in needed libraries 
```{r, include=FALSE}
library(tidyverse)
library(RColorBrewer)
library(ggplot2)
#library(haven)
#library(plotly)
#install.packages('rgl')
#library(rgl)
library(ISLR)
```

## 2. File management
### Create variables for directories
```{r file_management, include=FALSE}
project.dir <- getwd() #naeem
output.dir <- "/Output"
data.dir <- "C:/Users/Naeem Cho/Desktop/School Work/Statistical Inference/Datasets"
setwd(project.dir)
getwd()
```

## 3. Importing Data
```{r, echo = FALSE, include=FALSE}
advertising <- read_csv(file.path(data.dir, "Advertising.csv"))
```


# Problem #1

We will work on the familiar _Advertising_ data set.

  1. First, let's try using just the _TV_ budget to predict the _Sales_.
    a. Fit the $Sales \sim TV$ regression model, provde the residuals-vs-fitted plot. Is there an issue? If yes, how do we fix it? Proceed to modify the model in order to fix it.
    b. Having fixed the issue from part 1(a), check the residuals-vs-fitted plot for modified model. Are there any other issues with the plot now? If yes, proceed to modify the model in order to fix it (__Hint__: polynomial regression).
    c. For your finalized model from part 1(b), provide the QQ-norm plot. Does the fit look normal? If not quite, is it that big of a deal for inference purposes? Why?
    d. For your finalized model from part 1(b), provide the plot of the data along with fitted line/curve (make sure to __enumerate__ the points using _text()_ function, like I did with _us_statewide_crime_ in Slides 5.R) Indicate observations that appear to be: 1) regression outliers; 2) high-leverage; 3) influential outliers.
    e. From part 1(d), pick the observation you believe to be the most influential outlier. Proceed to fit the regression 1) __with__ and 2) __without__ that observation included. Report the __fitted equations, RSE, and__ $R^2$ values, __comment on differences observed. DON'T INCLUDE THE OUTPUT OF $R$ COMMANDS__ (eval = F).
    f. What do you think we should do with that observation? Why?