---
title: "Stats. Inference Assignment 4"
author: "Naeem Chowdhury"
date: "5/25/2020"
output: pdf_document
---


##1. Setup
### options
Set up global options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70))
knitr::opts_chunk$set(fig.height=4, fig.width=6)
```

### libraries
Load in needed libraries 
```{r, include=FALSE}
library(tidyverse)
library(RColorBrewer)
library(haven)
```

## 2. File management
### Create variables for directories
```{r file_management, include=FALSE}
project.dir <- getwd() #naeem
output.dir <- "/Output"
data.dir <- "C:/Users/Naeem Cho/Desktop/School Work/Statistical Inference/Datasets"
setwd(project.dir)
getwd()
```

## 3. Importing Data
```{r}
cpu.data <- read_csv(file.path(data.dir, "Intel_CPUs.csv"))
```


**The goal of this homework is to have you 1) Familiarized with the $\chi^2$ test of independence for contingency tables; 2) measuring practical strength of dependence between categorical variables; 3) familiarized with permutation test for contingency tables; 4) practice your R coding.**

# Problem #1

## 1. For all three examples on slide #24, proceed to

### a. Use the _my.chisq.test()_ function you've defined in the previous HW in order to confirm the $X^2$ and $p$-values. __Hint:__ Make sure to convert the %'s into counts first.

### b. Calculate the difference in proportion between males and females that attend religious services weekly. Calculate the risk ratio between males and females that attend religious services weekly.

### c. Based on your answers to parts (a) - (b), as $n$ increases, what do you notice with respoect to statistical significance? Practical significance?

## 2. Do exercises **11.32** and **11.33** from the **Agresti** book.

# Problem #2

## 1. Code up your own _my.permutation.test()_ function to conduct permutation tests on contingency tables. As inputs, it should take a dataframe with two categorical variables as columns (first one - explanatory, second one - response), and the number of randomly generated permutationts to be executed. As output, it should provide:

    - contingency table for the data frame,
    
    - permutation $p$-value,
    
    - plot the histogram of permutation distribution for $X^2$ statistic.
    
  To obtain $X^2$ values for each permutation, you are allowed to use $R$'s _chisq.test()_ function.
  
## 2. Proceed to apply the _my.permutation.test()_ function (and subsequently interpret the results) to:

### a. GPU_data, with 10,000 permutations. What's the conclusion? Compare the resulting histogram with the one in the slides (they should be roughly similar).

### b. Intel_CPU data, with just 1,000 permutations. What's the conclusion? Compare the shape of the resulting histogram with the density of $\chi^2$ distribution with appropriate degrees of freedom. What does it tell us about whether $\chi^2$-test results from HW3 were appropriate for Intel_CPU data?