---
title: 'Stats. Inference Homework #2'
author: "Naeem Chowdhury"
date: "2/13/2020"
output:
  html_document:
    df_print: paged
---

## 9.66 Exploring Type II Errors

Refer to the web app from Activity 2 at the end of this section, now assuming that we are using the two-sided test $H_0: p = 0.33$ against $H_a: p \neq 0.33$.

  a. Explain the effect of increasing the sample size on the probability of a Type II error when the true $p = 0.50$.
  
### Solution a.
  
  As sample size increases, the probability of a Type II error decreases. This is because as sample size increases, value of the null hypothesis $p_0$ falls increasingly  more standard deviations from the mean.
  
  b. Use the app to find the sample size needed to achieve a power of at least 90% when truly $p = 0.5$.
  
### Solution b.
  
  A sample size of 84 is suffcient to achieve a power of at least 90%.
  
  c. For a fixed sample size, do you think the probability of a Type II error will increase or decrease when the true $p$ is 0.40 instead of 0.50? Check your answer with the app. 
  
### Solution c. 

  I think that for a fixed sample size, the probability of a Type II error will increase when the true $p$ is 0.40 instead of 0.50. This is because the center of the $H_a$ distribution will be closer to the value $p_0$ of the null hypothesis.
  
  Indeed, it is true!
  
  d. Does the power increase or decrease when the significance level is 0.10 instead of 0.05? Check your answer with the app.
  
### Solution d.

  The power increases when the significance level is 0.10 instead of 0.05. This is because by increasing our significance level, we are decreasing the radius around $p_0$ under which we compute the area under the $H_a$ distribution. This decreases the Type II error probability, increasing the Power.
  
  
## Problem 2.

  Presume you've developed a skin cancer treatment and you were granted permission to test it out on patients. You would like to test if its accuracy differs from the golden standard method which has 28% cure rate. In particular, you'd want the ability to correctly detect a difference of 10%. 

  a. Formulate the hypotheses for the one-sample proportion test.
  
### Solution a.

  Let $p$ be the proportion of patients cured under the skin cancer treatment that I developed.

  Then the hypotheses are...

  $$H_0: p = 0.28.$$

  $$H_a: p \neq -0.28.$$


  b. Interpret the statement: "At a 0.05 significance level, the significance test will have 0.77 power when detecting a difference of 10%."
  
### Solution b.

  With a confidence level of 95%, we can detect a 10% difference between our drug and the leading drug  77% of the time.
  

  c. Presume we witness your treatment's results for 100 patients. Obtain the power of one-sample proportion test at $\alpha = 0.05$ significance level when detecting a difference of 10%. 
  
 
  
```{r}
n <- 100
p0 <- 0.28
pA <- 0.38
alpha <- 0.05

reject <- qnorm(1-alpha/2, mean=p0, sd= sqrt(p0*(1-p0)/n))
# Calculating the Type II error, the probability that we fail to observe a difference when there actually is one
type_II <- pnorm(reject, mean = pA, sd = sqrt(p0*(1-p0)/n)) - pnorm(-reject, mean = pA, sd = sqrt(p0*(1-p0)/n))
# Power = 1 - type_II
powah = 1 - type_II
```

  
  d. Presume we witness your treatment's results for 100 patients. Obtain the $\alpha$ significance level needed for your test to have the power of $0.85$ when detecting a difference of 10%.
  
   To solve this question, we can utilize a function I wrote for the last assignment.



  
```{r}
n <- 100
p0 <- 0.28
pA <- 0.38
alpha <- 0.235

# Quantiles
reject <- qnorm(1-alpha/2, mean=p0, sd= sqrt(p0*(1-p0)/n))

type_II <- pnorm(reject, mean = pA, sd = sqrt(p0*(1-p0)/n)) - pnorm(-reject, mean = pA, sd = sqrt(p0*(1-p0)/n))

powah <- 1 - type_II

powah
```

  
  e. What \# of patients is needed for your test to have the power of 0.85 when detecting a difference of 10%?
  
  
  Based on the previous problem, our confidence level would be $1 - \alpha = 0-.765$ to have the desired power. Utilizing a function from the previous homework...
  
```{r}
prop.sample.size <- function(m, conf) {
  alpha = 1-conf
  n <- (qnorm(1-(alpha/2))^2 * 0.5^2)/(m^2)
  response <- c("The sample size for a one-sample proportion test with a desired margin of error", m, " and given confidence level", conf, " is ", n)
  return(response)
}

prop.sample.size(0.1, 0.735)

```
  
  So we would only need about 31 participants to reach the desired power.
  
  
  f. In this case, do you think it is more important for your test to have lower significance level or higher power? Explain.
  
  If we lower our significance level, we might believe that our treatment is better when in fact the golden treatment performs better. Having a higher power lets us know that we are less likely to commit the error that in accepting there is no important difference between the two errors when it actually is effective. Power is more important than significance level because it encourages the development of our drug, where as having a lower significance level might cause us to waste millions of dollars on a drug that is no better than what we already have.
  

  g. Provided that you don't have the resources to recruit more than $n = 100$ patients, what can you do in order to increase the power of your test for detecting 10% difference.
  
  To increase the power of our test, we can decrease the size of our confidence interval. 

