---
title: "Stats. Inference Assignment 3"
author: "Naeem Chowdhury"
date: "5/19/2020"
output: pdf_document
---


##1. Setup
### options
Set up global options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70))
knitr::opts_chunk$set(fig.height=4, fig.width=6)
```

### libraries
Load in needed libraries 
```{r}
library(tidyverse)
library(RColorBrewer)
```

## 2. File management
### Create variables for directories
```{r file_management}
project.dir <- getwd() #naeem
output.dir <- "/Output"
data.dir <- "C:/Users/Naeem Cho/Desktop/School Work/Statistical Inference/Datasets"
setwd(project.dir)
getwd()
```

# Problem #1

## 1. Code up your own _my.chisq.test()_ function that will perform a $\chi^2$ test. As a single argument, it should just take a contingency table of arbitrary size. As output, it should provide:

- Calculated $\chi^2$ statistic
- $p$-value

Caclulating the expected cell counts under $H_0$ hypothesis should constitute a critical part of your function definition. Don't use neither _chisq.test()_ nor _prop.test()_, nor any other "fancy cheat" built-in functions inside your function's definition.

```{r}
data <- matrix(c(18, 20, 15, 15, 10, 55, 65, 70, 30), nrow=3)
data

my.chisq.test <- function(data) {
  
  
  
}
```





