---
title: '2020-04-11 HW8'
author: "Naeem Chowdhury"
date: "04/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setup

### Libraries
Load in needed libraries 
```{r libraries}
library(tidyverse) # For data processing and 

spot_color = "#9ecae1" 
text_color = "#525252"
```


### Data
```{r}
# Store full Auto.csv dataset from ISLR, not including 'name', as requested.
auto_df <- read_csv('../../Datasets/Auto.csv') %>% select(-name)
# Store full Advertising.csv dataset from ISLR
advertising_df <- read_csv('../../Datasets/Advertising.csv')
```

## Problem #1

1. For _Auto_ data set from ISLR library. Proceed to conduct variable selection via backward AIC approach. Which _R_ function allows us to do that? Which variable(s) ended up being dropped from the model? Proceed to:
  a. Explain what is meant by "Df", "Sum of Sq", "RSS" and "AIC" in the tables outputted by _step()_ function.
  b. Explain why the aglorithm stopped (!) on that particular subset of variables. _Hint_: What does "<none>" represent? Why is it of interest?)
  c. Provide the formula of the fitted model with remaining variables.
  
  